---
title: '`Quiz 4` Getting and Cleaning Data'
author: '`r if(knitr::is_html_output()) {"&#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Anderson H Uyekita"} else {NULL}`'
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
      smooth_scroll: true
  github_document: default
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache=TRUE,message=FALSE,warning=FALSE)
library(tidyverse)
library(magrittr)
library(xml2)
library(data.table)
library(httr)
library(sqldf)
library(DT)
library(lubridate)
library(rmarkdown)
```

`r if(!knitr::is_html_output()) {"* &#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Author: Anderson H Uyekita"}`
* &#x1f4da; Specialization: [Data Science: Foundations using R Specialization](https://www.coursera.org/specializations/data-science-foundations-r){target='_blank' rel='noopener'}
* &#x1f4d6; Course: [Getting and Cleaning Data](https://www.coursera.org/learn/data-cleaning){target='_blank' rel='noopener'}
    * &#x1F9D1;&#x200d;&#x1F3EB; Instructor: Jeffrey Leek
* &#x1F4C6; Week 4
    * Start: 2022/05/23
    * Finish: 2022/05/24
* `r if(!knitr::is_html_output()) {"&#x1F30E; Rpubs: [Interactive Document](https://rpubs.com/AndersonUyekita/quiz-4_getting-and-cleaning-data)"}else{"&#x1F4E6; Github Repository: [Static Document](https://github.com/AndersonUyekita/getting-and-cleaning-data/blob/main/Week%204/getting-and-cleaning-data_quiz-4.md){target='_blank' rel='noopener'}"}`

***

```{r creating_folder,cache=TRUE}
# Checking if the subfolder already exists.
if (!dir.exists("data")) {
    
    # Creating a subfolder to store the data.
    dir.create(path = "./data")
}
```

## Question 1

```{r,cache=TRUE}
# Downloading the file "The American Community Survey".
utils::download.file(url = "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv",
                     destfile = "./data/acs.csv",
                     mode = "wb")

# Downloading the code book.
utils::download.file(url = "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FPUMSDataDict06.pdf",
                     destfile = "./data/code_book.pdf",
                     mode = "wb")
```

Importing the downloaded file into R Object.

```{r,cache=TRUE}
# Loading the ACS data.
acs <- utils::read.csv(file = "./data/acs.csv")

# Performing the given expression.
strsplit(x = colnames(acs), "wgtp")[123]
```

## Question 2

```{r,cache=TRUE}
# Downloading the GDP data.
utils::download.file(url = "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FGDP.csv",
                     destfile = "./data/gdp.csv")

# Loading GDP data.
df_gdp <- utils::read.csv(file = "./data/gdp.csv",
                          skip = 3,                   # The first rows of this database is blank
                          header = TRUE)              # Forcing the read.csv readh the first rows as header.

# First rows of GDP data.
    # CASE: github_document
    if(!knitr::is_html_output()) {
    
        # Static table using Kable Package.
        df_gdp %>%
            head() %>%
            kableExtra::kbl() %>%
            kableExtra::kable_styling()
    
    # CASE: hmtl_document
    } else {
    
        # Interactive table using DT package.
        DT::datatable(df_gdp)
    }

# Data Cleaning
df_gdp_tidy <- df_gdp %>%
    mutate(US.dollars. = sub(pattern = " ", replacement = "", x = US.dollars.)) %>%
    mutate(US.dollars. = gsub(pattern = ",", replacement = "", x = US.dollars.)) %>%
    mutate(gdp = as.numeric(US.dollars.)) %>%
    mutate(ranking = as.numeric(Ranking)) %>%
    select(-c(X.1, X.2, X.3, X.4, X.5, X.6)) %>%
    na.omit() %>%
    rename("CountryName" = Economy)
```

After the data manipulation, I have calculated the GDP average as follows:

```{r}
# Calculating the GDP Average
mean(df_gdp_tidy$gdp, na.rm = TRUE)
```

## Question 3

The pattern used to filter countries that start with "United" is: `^United`.

```{r,cache=TRUE}
# Readings: https://r4ds.had.co.nz/strings.html
#    ^ to match the start of the string.
#
# Solution 1:
grep(pattern = "^United", x = df_gdp_tidy$CountryName)

# Solution 2:
df_gdp_tidy %>% filter(grepl(pattern = "^United", x = CountryName))
```

Both solutions have the same 3 countries, which starts with the United.

## Question 4

```{r,cache=TRUE}
# Downloading the GDP data.
utils::download.file(url = "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FGDP.csv",
                     destfile = "./data/gdp.csv")

# Downloading the Educational data.
utils::download.file(url = "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FEDSTATS_Country.csv",
                     destfile = "./data/education.csv")

# Loading Education data.
df_education <- utils::read.csv(file = "./data/education.csv")

# First rows os Education Data.
    # CASE: github_document
    if(!knitr::is_html_output()) {
    
        # Static table using Kable Package.
        df_education %>%
            head() %>%
            kableExtra::kbl() %>%
            kableExtra::kable_styling()
    
    # CASE: hmtl_document
    } else {
    
        # Interactive table using DT package.
        DT::datatable(df_education)
    }

# Loading GDP data.
df_gdp <- utils::read.csv(file = "./data/gdp.csv",
                          skip = 3,                   # The first rows of this database is blank
                          header = TRUE)              # Forcing the read.csv readh the first rows as header.

# CLEANING
df_gdp <- df_gdp %>%
    mutate(US.dollars. = sub(pattern = " ", replacement = "", x = US.dollars.)) %>%
    mutate(US.dollars. = gsub(pattern = ",", replacement = "", x = US.dollars.)) %>%
    mutate(US.dollars. = as.numeric(US.dollars.)) %>%
    mutate(Ranking = as.numeric(Ranking))

# CLEANING
df_gdp_tidy <- df_gdp %>%
    select(-c(X.1, X.2, X.3, X.4, X.5, X.6)) %>%
    rename(country_shortcode = X, ranking = Ranking, country = Economy, gdp = US.dollars.) %>%
    mutate(gdp = as.numeric(gdp)) %>%
    na.omit() %>%
    filter(ranking %in% 1:190)

# CLEANING
df_education_tidy <- df_education %>%
    rename(country_shortcode = CountryCode)

# Merging
df_gdp_edu <- merge(x = df_education_tidy, y = df_gdp_tidy) 
```

In the `Special.Notes` column, it is possible to track when the Fiscal Year ends. I have used the expression "June 30" to find all the countries with their fiscal year ending on this date.

```{r}
# Printing the countries with the fiscal year ending on 30 June (end of June).
    # CASE: github_document
    if(!knitr::is_html_output()) {
    
        # Static table using Kable Package.
        df_gdp_edu %>%
            select(country, Special.Notes) %>%
            filter(grepl(pattern = "June 30", x = Special.Notes)) %>%
            kableExtra::kbl() %>% kableExtra::kable_styling()
    
    # CASE: hmtl_document
    } else {
    
        # Interactive table using DT package.
        df_gdp_edu %>%
            select(country, Special.Notes) %>%
            filter(grepl(pattern = "June 30", x = Special.Notes)) %>%
            DT::datatable()
    }
```

Counting the number of countries.

```{r}
# Filtering and counting the countries with fiscal year ending on June.
df_gdp_edu %>%
    select(country, Special.Notes) %>%
    filter(grepl(pattern = "June 30", x = Special.Notes)) %>%
    nrow()
```

## Question 5

Executing the given code:

```{r,cache=TRUE,include=TRUE}
# Loading the quantmod package
library(quantmod)

# Gathering data from Amazon.
amzn = getSymbols("AMZN",auto.assign=FALSE)

# Gathering the dates from each data.
sampleTimes = index(amzn)
```

I have converted it into a tibble dataframe to be able to use the tidyverse package.

```{r}
# Converting the amzn object into Tibble.
data_q5 <- as_tibble(amzn)

# Adding the date column.
data_q5['date'] <- sampleTimes

# Filtering data only from 2012.
data_q5_2012 <- data_q5 %>%
    filter(date >= "2012-01-01" & date <= "2012-12-31")

# Showing the data gathered.
    # CASE: github_document
    if(!knitr::is_html_output()) {
    
        # Static table using Kable Package.
        data_q5_2012 %>%
            head() %>%
            kableExtra::kbl() %>%
            kableExtra::kable_styling()
    
    # CASE: hmtl_document
    } else {
    
        # Interactive table using DT package.
        DT::datatable(data_q5_2012)
    }
```

The number of observations in 2012:

```{r}
# How many values in 2012
nrow(data_q5_2012)
```

I have used the Lubridate package to work with a date type. In addition, I have added a column to store the wee day.

```{r}
# Ensuring to show the Week days in English.s
Sys.setlocale("LC_ALL","English")

# Adding a new column to store the wee days.
data_q5_2012_wdays <- data_q5_2012 %>%
    mutate(wday = lubridate::wday(x = lubridate::ymd(date), label=TRUE))

# Showing the data with the new column of week day.
    # CASE: github_document
    if(!knitr::is_html_output()) {
    
        # Static table using Kable Package.
        data_q5_2012_wdays %>%
            head() %>%
            kableExtra::kbl() %>%
            kableExtra::kable_styling()
    
    # CASE: hmtl_document
    } else {
    
        # Interactive table using DT package.
        DT::datatable(data_q5_2012_wdays)
    }
```

Finally, the number of Monday in 2012:

```{r}
# Number of observations in 2012 on Mondays.
data_q5_2012_wdays%>%
    filter(wday == "Mon") %>%
    nrow()
```