---
title: '`Quiz 3` Getting and Cleaning Data'
author: "Anderson H. Uyekita"
output:
  github_document: default
  html_document:
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
      smooth_scroll: true
always_allow_html: true
---

* Specialization: [Data Science: Foundations using R Specialization](https://www.coursera.org/specializations/data-science-foundations-r)
* Course: [Getting and Cleaning Data](https://www.coursera.org/learn/data-cleaning)
    * Instructor: Jeffrey Leek
* Week 3
    * Start: 2022/05/23
    * Finish: 2022/05/23
    
***

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache=TRUE,message=FALSE,warning=FALSE)
library(tidyverse)
library(magrittr)
library(xml2)
library(data.table)
library(httr)
library(sqldf)
library(data.table)
library(kableExtra)
```

```{r creating_folder,cache=TRUE}
# Checking if the subfolder already exists.
if (!dir.exists("data")) {
    
    # Creating a subfolder to store the data.
    dir.create(path = "./data")
}
```

## Question 1

```{r,cache=TRUE}
# Downloading the American Community Survey file.
utils::download.file(url = "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv",
                     destfile = "./data/acs.csv")

# Downloading the code book.
utils::download.file(url = "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FPUMSDataDict06.pdf",
                     destfile = "./data/code_book.pdf",
                     mode = "wb")

# Loading CSV file into R object.
acs <- utils::read.csv(file = "./data/acs.csv")
```

The American Community Survey Dataset.

```{r,cache=TRUE}
# First rows of ACS and first columns.
acs %>% select(c(1:5)) %>%
    head(5) %>%
    kableExtra::kbl() %>%
    kableExtra::kable_styling()
```

Dimensions of ACS dataset:

```{r}
# Number of observations and Variables
dim(acs)
```

Following the question instruction, I will only display the first 3 values.

Based on the `code book` provided:

* *households on greater than 10 acres*, means: ACR == 3
* *sold more than $10,000 worth of agriculture products*, means: AGS == 6

```{r}
# According to the statement.
agricultureLogical <- acs$ACR == 3 & acs$AGS == 6

# Performing the given expression:
head(which(agricultureLogical), 3)
```

## Question 2

```{r,cache=TRUE}
# Loading the JPEG package.
library(jpeg)

# Downloading the KPEG file.
utils::download.file(url = "https://d396qusza40orc.cloudfront.net/getdata%2Fjeff.jpg",
                     destfile = "./data/image_file.jpeg", mode = "wb")

# Loading the JPEG file. Following the instruction to set NATIVE as TRUE.
image <- jpeg::readJPEG(source = "./data/image_file.jpeg",
                        native = TRUE)

# Answering the quantiles 30% and 80%.
c(quantile(x = image, 0.3), quantile(x = image, 0.8))
```

My answer is a bit different, as warned by the question instruction.

## Question 3

```{r}
# Downloading the GDP data.
utils::download.file(url = "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FGDP.csv",
                     destfile = "./data/gdp.csv")

# Downloading the Educational data.
utils::download.file(url = "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FEDSTATS_Country.csv",
                     destfile = "./data/education.csv")

# Loading Education data.
df_education <- utils::read.csv(file = "./data/education.csv")

# First rows os Education Data.
df_education %>%
    select(c(1:6)) %>%
    head() %>%
    kableExtra::kbl() %>%
    kableExtra::kable_styling()

# Loading GDP data.
df_gdp <- utils::read.csv(file = "./data/gdp.csv",
                          skip = 3,                   # The first rows of this database is blank
                          header = TRUE)              # Forcing the read.csv readh the first rows as header.

# First rows of GDP data.
df_gdp %>%
    head() %>%
    kableExtra::kbl() %>%
    kableExtra::kable_styling()
```

We need to match those databases on the `country shortcode`. For this reason, we need to be sure if this `country shortcode` is present in both datasets. The GDP database is not cleaned.

As you can see, we do not have columns with identical names neither GDP columns with numeric.

```{r}
# DATA CLEANING: Fixing the GDP to be number and removing comas.
df_gdp <- df_gdp %>%
    mutate(US.dollars. = sub(pattern = " ", replacement = "", x = US.dollars.)) %>%
    mutate(US.dollars. = gsub(pattern = ",", replacement = "", x = US.dollars.)) %>%
    mutate(US.dollars. = as.numeric(US.dollars.)) %>%
    mutate(Ranking = as.numeric(Ranking))

# DATA CLEANING: Removing unecessary columns of GDP database.
df_gdp_tidy <- df_gdp %>%
    select(-c(X.1, X.2, X.3, X.4, X.5, X.6)) %>%
    rename(country_shortcode = X, ranking = Ranking, country = Economy, gdp = US.dollars.) %>%
    mutate(gdp = as.numeric(gdp)) %>%
    na.omit() %>%
    filter(ranking %in% 1:190)

# DATA CLEANING: Using the same column name to Country.
df_education_tidy <- df_education %>%
    rename(country_shortcode = CountryCode)
```

After the cleaning process, I merged those datasets into one.

The 13th country in the ordered dataset in a crescent way:

```{r}
# Merging Education and GDP databses.
df_gdp_edu <- merge(x = df_education_tidy, y = df_gdp_tidy) 

# Searching the 13th country in crescent order.
df_gdp_edu%>%
    select(country_shortcode, Long.Name, gdp) %>%
    arrange(gdp) %>%
    filter(order(gdp) == 13)
```

The number of countries with GDP and Education database match. 

```{r}
# Counting the number of countries with match.
df_gdp_edu %>% nrow()
```

## Question 4

Investigating the `Income.Group` variable. I just want to see what is the "levels".

```{r}
# Classes of Income Group
unique(df_gdp_edu$Income.Group)
```

Now, I could calculate the average of each `Income.Group` level. The GDP average of `High income: OECD` and `High income: nonOECD` from `Income.Group` could be found below:

```{r}
# Calculating the average of each Income Group Level.
df_gdp_edu %>% group_by(Income.Group) %>% 
    summarise(AVG = mean(as.numeric(ranking)))
```

## Question 5

Let's see the quantiles boundaries.

```{r}
# Creating the groups limits.
quantile(x = as.numeric(df_gdp_edu$ranking), c(0.0, 0.2, 0.4, 0.6, 0.8, 1.0))
```

Following the question instructions, I need to create a CUT. To do so, I will create a new variable, `ranking_group`, to store these grouping classifications.

```{r}
# Adding a new column to the merged dataframe.
df_gdp_edu$ranking_group <- cut(x = df_gdp_edu$ranking, breaks = quantile(x = as.numeric(df_gdp_edu$ranking), c(0.0, 0.2, 0.4, 0.6, 0.8, 1.0)))

# Printing the results of this new column.
df_gdp_edu %>% select(country_shortcode, Long.Name,gdp,ranking_group) %>% head()
```

I have already inserted this new variable to classify each country based on those quantiles. I need to count the number of countries in `Lower middle income` in `Income.Group`.

```{r}
df_gdp_edu %>%
    group_by(ranking_group) %>%
    filter(Income.Group == "Lower middle income") %>%
    summarise(n = n())
```