---
title: '`Quiz 2` Getting and Cleaning Data'
author: '`r if(knitr::is_html_output()) {"&#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Anderson H Uyekita"} else {NULL}`'
output:
  github_document: default
  html_document:
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
      smooth_scroll: true
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache=TRUE,message=FALSE,warning=FALSE)
library(tidyverse)
library(magrittr)
library(xml2)
library(data.table)
library(httr)
library(sqldf)
library(rmarkdown)
```

`r if(!knitr::is_html_output()) {"* &#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Author: Anderson H Uyekita"}`
* &#x1f4da; Specialization: [Data Science: Foundations using R Specialization](https://www.coursera.org/specializations/data-science-foundations-r){target='_blank' rel='noopener'}
* &#x1f4d6; Course: [Getting and Cleaning Data](https://www.coursera.org/learn/data-cleaning){target='_blank' rel='noopener'}
    * &#x1F9D1;&#x200d;&#x1F3EB; Instructor: Jeffrey Leek
* &#x1F4C6; Week 2
    * &#x1F6A6; Start: 2022/05/21
    * &#x1F3C1; Finish: 2022/05/21
* `r if(!knitr::is_html_output()) {"&#x1F30E; Rpubs: [Interactive Document](https://rpubs.com/AndersonUyekita/quiz-2_getting-and-cleaning-data)"}else{"&#x1F4E6; Github Repository: [Static Document](https://github.com/AndersonUyekita/getting-and-cleaning-data/blob/main/Week%202/getting-and-cleaning-data_quiz-2.md){target='_blank' rel='noopener'}"}`

***

This Quiz is part of the Getting and Cleaning Data Course.

To turn the folder more structured, I will create a subfolder to host the dataset.

```{r creating_folder,cache=TRUE}
# Checking if the subfolder already exists.
if (!dir.exists("data")) {
    
    # Creating a subfolder to store the data.
    dir.create(path = "./data")
}
```

## Question 1

```{r}
# 1. Find OAuth settings for github:
#    http://developer.github.com/v3/oauth/
oauth_endpoints("github")

# 2. My credentials App.
#
# Go to https://github.com/settings/developers and create a "New OAuth App".
myapp <- oauth_app("github",
                   key = Sys.getenv("key"),        # key defined in .Renviron
                   secret = Sys.getenv("secret"))  # secret defined in .Renviron

# 3. Get OAuth credentials
github_token <- oauth2.0_token(endpoint = httr::oauth_endpoints("github"),
                               app = myapp)

# 4. Requesting the Github API to gather info from Jeff repositories.
request <- httr::GET(url = "https://api.github.com/users/jtleek/repos",
                     config = httr::config(token = github_token))

# Checking the GET. If status == 200 means success!
request$status_code

# Converting into R object from a JSON file.
df_req <- content(x = request, as = "parsed")

# Number of Repositories in Jeff Leek Github
length(df_req)

# List of repositories initialization
repos_names <- vector()
data_creation <- vector()

# Loop to gather the repo's names.
for (i in 1:length(df_req)) {
    repos_names <- append(repos_names, df_req[[i]]$name)
    data_creation <- append(data_creation, df_req[[i]]$created_at)
}

# Filtering the datasharing repository and its data creation.
data.frame(cbind(repos_names, data_creation)) %>% filter(repos_names =="datasharing")
```

## Question 2

```{r q2}
# Downloading the file
download.file(url = "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06pid.csv",
              destfile = "./data/american_community_survey.csv")

# Loading the CSV file.
acs <- read.csv(file = "./data/american_community_survey.csv")
```

How many rows and columns?

```{r}
# Dimensions of this dataset.
dim(acs)     # [ROWS] [COLUMNS]
```

Comparing the options:

```{r}
# Solving this problem using tidyverse.
method_tidyverse <- acs %>% filter(AGEP < 50) %>% select(pwgtp1)

# Solving using the SQL
method_sql <- sqldf("select pwgtp1 from acs where AGEP < 50")

# Checking if both are identical.
identical(method_tidyverse, method_sql)
```

## Question 3

```{r}
# Subsetting the ACS dataset
method_r <- unique(acs$AGEP)

# The unique(acs$AGEP) retuns a vector.
# The sqldf("select distinct AGEP from acs") returns a dataframe. For this reason,
# I have selected the only columns of its dataframe to be comparable.
identical(method_r,
          sqldf("select distinct AGEP from acs")$AGEP)
```

## Question 4

```{r}
# Open a connection to a website.
#
# ATENTION!! I have to add the "www" to the given URL.
#
soup <- url("https://www.biostat.jhsph.edu/~jleek/contact.html")

# Parse the content of the given CONNECTION in text.
soup_parse <- readLines(con = "https://www.biostat.jhsph.edu/~jleek/contact.html")

# Closing connection.
close(soup)

# Delving the data.
nchar(soup_parse[10])
nchar(soup_parse[20])
nchar(soup_parse[30])
nchar(soup_parse[100])
```

## Question 5

```{r}
# Downloading the .FOR file.
download.file(url = "https://d396qusza40orc.cloudfront.net/getdata%2Fwksst8110.for",
              destfile = "./data/data_q5.for")

# Readings: https://www.rdocumentation.org/packages/utils/versions/3.6.2/topics/read.fortran
# I have read this website to understand how the format works.

# Loading the file into R Object
data_q5 <- read.fortran(file = "./data/data_q5.for",
                        format = c("A10", "X5",
                                   "F4.0",
                                   "F4.0", "X5",
                                   "F4.0",
                                   "F4.0", "X5",
                                   "F4.0",
                                   "F4.0", "X5",
                                   "F4.0",
                                   "F4.0", "X5"),
                        skip = 4)

# Sum of the 4th and 9th columns.
sum(data_q5$V4 + data_q5$V9)
```